\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, amsthm, commath, enumitem, graphicx, nopageno, quoting}
\usepackage[margin=1cm]{geometry}

\title{Computer Science 452 - Homework Assignment \#3}
\author{Hari Amoor, NetID: hra25}
\date{February 11, 2020}

\begin{document}
\maketitle

\section*{Problem 1.2.1: Convert the following finite automata to regular expressions.}
\begin{enumerate}[label=(\alph*)]
	\item The given automaton computes the regular expression $ba^{*}$.
	\item The given automaton computes the regular expression $((a \cup b)a^{*}b(bb)^{*}a)^{*}$.
\end{enumerate}

\section*{Problem 1.29b, Claim: The language $A = \{www \mid w \in \{a, b\}^{*}\}$ is not regular.}
\begin{proof}
	Suppose for contradiction that $A$ is regular. Let $p$ be the Pumping length of $A$, and $w = a^{p}b^{p}$. Then, $x = www \in A$. \\
	\newline
	The string $x$ is sufficiently large, so the decomposition $x = qrs$ must exist as per the Pumping lemma. Since $\abs{qr} \leq p$ and $r \neq \varepsilon$, $r$ must consist only of the symbol "a". Then, the string $qr^{2}s$ has more "a"s than "b"s, so it cannot be in $A$. \\
	\newline
	This result contradicts the Pumping lemma; thus, it is true that $A$ is not regular.
\end{proof}

\section*{Problem 1.32, Claim: The language $B$ as defined is regular.}
\textbf{NOTE:} Arithmetic is done in $\mathbb{Z}_{2}$ when describing $\delta$.
\begin{proof}
	It suffices to show that $B^{R}$ is regular. We construct a DFA to recognize $B^{R}$. \\
	\newline
	We assume w.l.o.g. that $\varepsilon \in B^{R}$. The DFA $M_{B} = (Q, \sum_{3}, \delta, q_{0}, F)$, where:
	\begin{align}
		Q = \{q_{0}, q_{1}, q_{e}\}; \\
		F = \{q_{0}\}; \\
		\delta\left(q_{i}, \begin{bmatrix} a \\ b \\ c \end{bmatrix}\right) = \begin{cases} q_{a + b + i} & q_{i} \neq q_{e} \text{ and } a + b + i = c \\ q_{e} & \text{otherwise}\end{cases}
	\end{align}
	recognizes $B$. Thus, the claim holds.
\end{proof}

\section*{Problem 1.35, Claim: The language $E$ as given is not regular.}
\begin{proof}
  Suppose for contradiction that $E$ is regular. Let $p$ be the Pumping length of $E$, and choose $s = \begin{bmatrix} 0 \\ 1 \end{bmatrix}^{p} \begin{bmatrix} 1 \\ 0 \end{bmatrix}^{p}$. \\
  \newline
  By definition, $s \in E$. Furthermore, since $\abs{s} > p$, it holds that $s = xyz$ as per the Pumping lemma. That is, $xy$ must have length at most $p$. Therefore, it holds that $y$ only contains symbols $\begin{bmatrix} 0 \\ 1\end{bmatrix}$. \\
    \newline
    So, the number of symbols of the form $\begin{bmatrix} 0 \\ 1 \end{bmatrix}$ and those of the form $\begin{bmatrix} 1 \\ 0 \end{bmatrix}$ must be unbalanced in $xy^{2}z$, so $xy^{2}z \notin E$. This contradicts the Pumping lemma. Thus, the claim holds.
\end{proof}

\section*{Problem 1.40: Prove the following claims.}
\begin{enumerate}[label=(\alph*)]
\item \textbf{$NOPREFIX$ closes the set of regular languages.}
  \begin{proof}
    Let $M = (Q, \Sigma, \delta, q_{0}, F)$ be a finite-state machine, i.e. a DFA, that computes a regular language $A$. Define $M' = (Q, \Sigma, \delta', q_{0}, F)$ be a finite-state machine with $\delta'(q, a) = \delta(r, a)$ if $r \notin F$ and $\delta'(q, a) = \emptyset$ otherwise (remove all edges from the accepting states). Then, $M'$ computes the language $NOPREFIX(A)$. \\
    \newline
    Hence, $NOPREFIX(A)$ is regular, as required.
  \end{proof}

\item \textbf{$NOEXTEND$ closes the set of regular languages.}
  \begin{proof}
    Suppose $A$ is a regular language and that the DFA $M = (Q, \Sigma, \delta, q_{0}, F)$ recognizes $A$. Let $M' = (Q, \Sigma, \delta, q_{0}, F \setminus F')$, where $F'$ is the set of accepting states that have a path of length at least 1 to another accepting state. \\
    \newline
    Here, $M'$ recognizes $NOEXTEND(A)$, so $NOEXTEND(A)$ is regular, as required.
  \end{proof}
\end{enumerate}

\section*{Problem 1.4.6: Prove that the following languages are not regular.}
\begin{enumerate}[label=(\alph*)]
  \item \textbf{$L = \{0^{n}1^{m}0^{n} \mid m, n \geq 0\}$}
    \begin{proof}
      Suppose for contradiction that $L$ is regular. Let $p$ be the Pumping length of $L$, and $s = 0^{p}1^{p}0^{p} \in L$. Then, $s = xyz$ as per the Pumping lemma. \\
      \newline
      Now, since $\abs{xy} \leq p$, it holds that $y$ consists entirely of 0s. Then, $s' = xy^{2}z$ is of the form $0^{q}1^{p}0^{p}$, where $q > p$. Thus, it holds that $s' \notin L$. \\
      \newline
      This contradicts the Pumping lemma; thus, the claim holds.
    \end{proof}
  \item \textbf{$L = \{w \mid w \text{is not a palindrome}\}$}
    \begin{proof}
      It suffices to show that $L^{\complement}$ is not regular. Suppose for contradiction that $L^{\complement}$ is regular, and let $p$ be the Pumping length of $L^{\complement}$. \\
      \newline
      Take $s = aa^{R} \in L^{\complement}$ where $\abs{a} \geq p$ and $a^{R}$ is the reversal of $a$. We know that $s = xyz$ as per the Pumping lemma. Since $\abs{xy} \leq p$, we know that $xy$ is some prefix of $a$. So, $s' = xy^{2}z \notin L^{\complement}$. \\
      \newline
      This is a contradiction of the Pumping lemma. Thus, the claim holds.
    \end{proof}
  \item \textbf{$L = \{wtw \mid w, t \in \{0, 1\}^{+}\}$}
    \begin{proof}
      Suppose for contradiction that $L$ is regular, and let $p$ be the Pumping length of $L$. Fix $w = 0^{p}1^{p}$, and let $s = wtw \in L$ with $t \in L$ arbitrary. \\
      \newline
      We know that $s = xyz$ with $\abs{xy} \leq p$, as per the Pumping lemma, so $xy$ consists entirely of 0s. Thus, $s' = xy^{2}z$ has an unbalanced number of 0s; consequently, $s' \notin L$. \\
      \newline
      This contradicts the Pumping lemma; thus, the claim holds.
    \end{proof}
\end{enumerate}

\section*{Problem 1.49: Prove the following.}
\begin{enumerate}[label=(\alph*)]
\item \textbf{The language $B$ as given is regular.}
  \begin{proof}
    Let $k$ in the context of $B$ be arbitrary. We know there exists a regular language $L_{1}$ of strings of the form $1^{k}$, and that there exists a regular language $L{2}$ of strings $y$ s.t. $y$ contains at least $k$ instances of the symbol \"1\". Observe that $B$ is the concatenation of $L_{1}$ and $L_{2}$, which implies that $B$ is regular since regular languages are closed under concatenation.
  \end{proof}
\item \textbf{The language $C$ as given is not regular.}
  \begin{proof}
    Suppose for contradiction that $C$ is regular. Let $p$ be the Pumping length of $C$. \\
    \newline
    Let $s = 1^{p}y \in C$, and suppose that $y$ has $k \leq p$ instances of the symbol 1. We know that $s = xyz$ with $\abs{xy} \leq p$, as per the Pumping lemma, so $y$ consists of all 1s. \\
    \newline
    Now, consider $s' = xy^{n}z$ for $n > p$. It must be the case that $s' \notin C$. This contradicts the Pumping lemma; thus, the claim holds.
  \end{proof}
\end{enumerate}
\end{document}
